A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Figures

### 2020-11-21 writeUp/supplementary_materials.tex
added new ms file
updated dissection estimates with new gonotrophic estimates
updated SOM and MS to include higher kappa estimates
updated graph titles

\section{MRR experments}
Figures \ref{fig:mrr_lifeSpanVsRange} and \ref{fig:mrr_lifeSpanVsTrapDensity} show the estimates of lifespan versus the trapping range and spatial density of traps at study location. In these cases, the lifespans shown were obtained from the non-hierarchical model with the exponential survival model. The spatial density of traps was obtained assuming that the study area was a circular area of radius equal to the maximum trapping distance.

Fig. \ref{fig:mrr_female_blood_sugar} shows the lifespans obtained for each pre-release feeding category for each genus and across all studies using the hierarchical Bayesian model with exponential survival model. In the genus-level models, the feeding-specific effects (not pre-fed, blood-fed-only and sugar-fed-only) were estimated as independent random effects for each genus with standard normal priors. For the overall model, the feeding-specific effects were estimated as independent random effects across all studies, with a standard normal prior.

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
	\caption{\textbf{The estimated predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model. The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
	\label{fig:mrr_genusTopLevel}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trapping range.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:mrr_lifeSpanVsRange}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the trap density.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range. The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:mrr_lifeSpanVsTrapDensity}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
	\caption{\textbf{Posterior estimates of female mosquito mean lifespan according to pre-release feeding status across species, genus and overall groupings as determined from the MRR data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:mrr_female_blood_sugar}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for each study location and date.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_temperature}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
	\caption{\textbf{Posterior estimates of mean mosquito lifespan for each time-series versus the average monthly temperature for the four species with the most data.} The markers show the median posterior estimates, with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
\end{figure}

\section{Dissection analysis}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
	\caption{\textbf{Individual time-series estimates of adult mosquito mean lifespan ordered by species median as determined from the dissection data.} The middle line in each box shows the median estimates. The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:dissection_individualEstimates_allSpecies}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
	\caption{\textbf{Posterior estimates of female mosquito mean chronological lifespan across species, genus and overall groupings as determined from the dissection data.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:dissection_lifetimes_exponential_chron}
\end{figure}

\begin{table}[htbp!]
	\begin{tabular}{l|l|l|l|l|l|l|l|l}
		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
		\hline
		\textit{Anopheles} & \textit{sergentii} & 4.2 & 7.6 & 10.1 & 13.6 & 24.6 & 11.9 & 9.7 \\
		\textit{Anopheles} & \textit{gambiae s.l.} & 2.8 & 5.2 & 7.6 & 11.4 & 21.6 & 9.5 & 7.4 \\
		\textit{Culex} & \textit{thalassius} & 4.6 & 6.3 & 7.4 & 8.6 & 12.0 & 7.9 & 4.1 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 5.9 & 6.5 & 6.9 & 7.3 & 8.0 & 6.9 & 0.7 \\
		\textit{Anopheles} & \textit{minimus} & 3.1 & 5.2 & 6.9 & 9.0 & 16.4 & 8.1 & 9.4 \\
		\textit{Anopheles} & \textit{maculipennis} & 3.3 & 4.8 & 5.9 & 7.5 & 11.4 & 6.5 & 2.7 \\
		\textit{Culex} & \textit{pipiens} & 4.6 & 5.3 & 5.7 & 6.2 & 6.9 & 5.8 & 0.8 \\
		\textit{Mansonia} & \textit{uniformis} & 3.8 & 4.4 & 4.8 & 5.2 & 6.0 & 4.8 & 0.8 \\
		\textit{Anopheles} & \textit{rivulorum} & 3.4 & 4.0 & 4.5 & 4.9 & 5.9 & 4.5 & 0.8 \\
		\textit{Anopheles} & \textit{melas} & 3.3 & 3.9 & 4.3 & 4.8 & 5.5 & 4.4 & 0.7 \\
		\textit{Anopheles} & \textit{culicifacies} & 3.4 & 4.0 & 4.3 & 4.7 & 5.4 & 4.4 & 0.6 \\
		\textit{Aedes} & \textit{aegypti} & 2.1 & 3.5 & 4.3 & 5.0 & 7.9 & 5.0 & 11.5 \\
		\textit{Anopheles} & \textit{subpictus} & 2.6 & 3.6 & 4.3 & 5.1 & 8.3 & 5.0 & 7.7 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 2.9 & 3.6 & 4.1 & 4.7 & 5.6 & 4.2 & 0.9 \\
		\textit{Anopheles} & \textit{stephensi} & 2.4 & 3.4 & 4.1 & 4.9 & 6.9 & 4.3 & 1.8 \\
		\textit{Culex} & \textit{quinquefasciatus} & 1.7 & 2.8 & 4.0 & 5.5 & 9.0 & 4.5 & 2.5 \\
		\textit{Anopheles} & \textit{darlingi} & 2.7 & 3.5 & 3.9 & 4.5 & 5.5 & 4.0 & 1.0 \\
		\textit{Aedes} & \textit{polynesiensis} & 3.1 & 3.5 & 3.8 & 4.1 & 4.5 & 3.8 & 0.5 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 1.3 & 2.6 & 3.8 & 5.7 & 16.8 & 9.8 & 123.0 \\
		\textit{Culex} & \textit{annulirostris} & 2.3 & 3.1 & 3.7 & 4.4 & 5.9 & 3.9 & 1.8 \\
		\textit{Aedes} & \textit{samoanus} & 2.6 & 3.0 & 3.3 & 3.7 & 4.4 & 3.4 & 0.9 \\
		\textit{Aedes} & \textit{sollicitans} & 1.5 & 2.2 & 2.7 & 3.3 & 5.4 & 3.0 & 2.0 \\
		\textit{Aedes} & \textit{vexans} & 1.7 & 2.1 & 2.5 & 2.9 & 3.9 & 2.7 & 2.7 \\
		\textit{Anopheles} & \textit{cruzii} & 1.4 & 1.9 & 2.3 & 2.9 & 4.6 & 2.6 & 1.5 \\
		\textit{Anopheles} & \textit{bellator} & 1.2 & 1.7 & 2.0 & 2.5 & 4.3 & 2.5 & 7.2 \\
		\hline
		\textit{Anopheles} & \textit{} & 2.5 & 4.1 & 5.7 & 7.9 & 12.7 & 6.4 & 3.4 \\
		\textit{Mansonia} & \textit{} & 3.7 & 4.4 & 4.8 & 5.1 & 5.9 & 4.8 & 0.8 \\
		\textit{Culex} & \textit{} & 2.1 & 3.2 & 4.4 & 5.9 & 9.5 & 4.9 & 2.4 \\
		\textit{Aedes} & \textit{} & 2.4 & 3.1 & 3.5 & 3.9 & 4.7 & 3.5 & 0.7 \\
		\hline
		\textbf{Overall} & \textit{} & 2.2 & 3.5 & 5.1 & 6.8 & 10.7 & 5.5 & 2.7 \\
	\end{tabular}
	\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from dissection studies.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std. dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
	\label{tab:dissection_estimated_lifespans_chron}%
\end{table}

\subsection{Comparison of LBLs between the MRR and dissection analyses}
To conduct a comparison between the LBLs from the MRR analysis and from the dissections, we converted the latter lifespans from gonotrophic cycles to chronological age as described in Section \ref{sec:dissection_conversion}. We conducted ANOVA using the individual time series estimates of LBL for each of the 10 species with estimates from both MRR- and dissection-based analysis. The results of this analysis are shown in Table \ref{tab:comparison}.

\begin{table}[htbp]
	\centering
	\begin{tabular}{l|l|l|l|l}
		species & \multicolumn{1}{l}{$\text{df}_1$} & \multicolumn{1}{l}{$\text{df}_2$} & \multicolumn{1}{l}{F} & \multicolumn{1}{l}{p} \\
		\midrule
		\textit{Ae. aegypti} & 1     & 55    & 1.9   & 0.17 \\
		\textit{A. culicifacies s.l.} & 1     & 19    & 2.2   & 0.15 \\
		\textit{A. darlingi} & 1     & 6     & 80.7   & 0.00 \\
		\textit{An. farauti s.l.} & 1     & 14    & 2.1  & 0.17 \\
		\textit{A. gambiae s.l.} & 1     & 37    & 4.3   & 0.05 \\
		\textit{Cx. pipiens} & 1     & 13    & 0.4   & 0.56 \\
		\textit{A. quadrimaculatus} & 1     & 8     & 2.2     & 0.18 \\
		\textit{Cx. quinquefasciatus} & 1     & 13    & 2.2    & 0.16 \\
		\textit{A. sergentii} & 1     & 8     & 1.0    & 0.36 \\
		\textit{A. subpictus} & 1     & 2    & 11.4   & 0.08 \\
		\textit{Cx. tritaeniorhynchus} & 1     & 5     & 3.6   & 0.12 \\
		\textit{Ae. vexans} & 1     & 2     & 1.1   & 0.41 \\
	\end{tabular}%
	\caption{\textbf{The results of comparisons between LBLs for those species with data for MRRs and dissection-based studies.} The ANOVA tests were conducted using median posterior LBLs estimated at the time-series level.}
	\label{tab:comparison}%
\end{table}%

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
	\caption{\textbf{The statistical power to detect senescence for A. three case study populations as a function of B. study length and C. number released.} For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size. The error bars show the standard deviation in the power for each parameter set. The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
	\label{fig:mrr_mcPowerAnalysis_senescence}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/elpd_comparison.pdf}}
	\caption{\textbf{Comparing the evidence for age-dependent mortality by fitting the Gompertz model to data from the MRR versus dissection studies for the 12 species with data from both.} The points indicate the estimated predictive accuracy compared with the exponential model (see SOM Methods) and the lower and upper whiskers represent 2.5\% and 97.5\% quantiles on this quantity.}\label{fig:elpd_comparison}
\end{figure}

>>>>

\section{MRR experments}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_genus_vs_species_vs_top.pdf}}
	\caption{\textbf{MRR: predictive accuracy of the species- and overall-level models relative to the genus-level model.} The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model. The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy.}
	\label{fig:mrr_genusTopLevel}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsRange.pdf}}
	\caption{\textbf{MRR: mean mosquito lifespan versus trapping range.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:mrr_lifeSpanVsRange}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespanVsTrapDensity.pdf}}
	\caption{\textbf{MRR: mean mosquito lifespan versus trapping density.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The trap density was estimated assuming traps were contained within a circular area with radius equal to the trapping range. The black line shows a linear regression line estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:mrr_lifeSpanVsTrapDensity}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_female_blood_sugar.pdf}}
	\caption{\textbf{MRR: female mosquito mean lifespan by pre-release feeding status.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:mrr_female_blood_sugar}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifeSpanVsTemperature.pdf}}
	\caption{\textbf{MRR: mean mosquito lifespan versus average monthly air temperature.} The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_temperature}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_ThreeSpeciesVersusTemperature.pdf}}
	\caption{\textbf{MRR: mean mosquito lifespan versus average monthly air temperature by species.} Here, we include only those four species with the most data. The markers show the median posterior lifespan estimates for each time series with the lower and upper bounds indicating the 25\% and 75\% quantiles, respectively. The black line shows a regression line with linear and quadratic terms estimated using the median posterior lifetimes, with the grey shading indicating 95\% confidence intervals. All estimates of lifespan were obtained using the non-hierarchical exponential survival model.}\label{fig:mrr_ThreeSpeciesVersusTemperature}
\end{figure}

\section{Polovodova dissection analysis}

\begin{figure}[h]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_individualEstimates_allSpecies.pdf}}
	\caption{\textbf{Polovodova dissection: individual time-series estimates of adult mosquito mean lifespan.} The middle line in each box shows the median estimates. The left and right box whiskers show the 25\%, and 75\% posterior quantiles respectively. All estimates were obtained using the non-hierarchical exponential survival model.}
	\label{fig:dissection_individualEstimates_allSpecies}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifetimes_exponential_chron.pdf}}
	\caption{\textbf{Polovodova dissection: female mosquito mean chronological lifespan estimates.} The lifespans shown are for mosquitoes that were not fed with sugar or blood (for females) before release. The middle line in each box shows the median estimates and the solid dot indicates the mean. The left and right box edges show the 25\%, and 75\% posterior quantiles respectively. The whiskers show the range of the data, excluding points lying more than 1.5 times the interquartile range away from each edge of the box. The numbers before the start of the left whisker indicate the number of individual time-series within each species. All estimates were obtained using the hierarchical exponential survival model.}
	\label{fig:dissection_lifetimes_exponential_chron}
\end{figure}

\begin{table}[htbp!]
	\begin{tabular}{l|l|l|l|l|l|l|l|l}
		\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} & \textbf{75$\%$} & \textbf{95\%} &  \textbf{Mean} & \textbf{Std. dev.}\\
		\hline
	\textit{Culex} & \textit{thalassius} & 5.7 & 8.2 & 9.4 & 11.0 & 15.6 & 10.1 & 5.2 \\
	\textit{Anopheles} & \textit{sergentii} & 3.7 & 6.4 & 8.5 & 11.5 & 20.6 & 10.1 & 8.1 \\
	\textit{Culex} & \textit{pipiens} & 5.7 & 6.6 & 7.2 & 7.8 & 8.9 & 7.2 & 1.0 \\
	\textit{Anopheles} & \textit{gambiae s.l.} & 2.4 & 4.3 & 6.4 & 9.8 & 18.3 & 8.0 & 6.3 \\
	\textit{Anopheles} & \textit{farauti s.l.} & 4.8 & 5.5 & 5.9 & 6.3 & 6.9 & 5.9 & 0.6 \\
	\textit{Anopheles} & \textit{minimus} & 2.7 & 4.4 & 5.8 & 7.7 & 13.7 & 6.9 & 8.1 \\
	\textit{Anopheles} & \textit{maculipennis} & 2.8 & 4.0 & 5.1 & 6.4 & 9.6 & 5.5 & 2.3 \\
	\textit{Culex} & \textit{quinquefasciatus} & 2.0 & 3.4 & 4.8 & 6.9 & 11.7 & 5.6 & 3.3 \\
	\textit{Aedes} & \textit{aegypti} & 2.3 & 3.7 & 4.7 & 5.6 & 9.3 & 5.6 & 12.3 \\
	\textit{Culex} & \textit{annulirostris} & 2.7 & 3.9 & 4.6 & 5.4 & 7.5 & 4.9 & 2.4 \\
	\textit{Culex} & \textit{tritaeniorhynchus} & 1.6 & 3.2 & 4.6 & 7.2 & 22.8 & 13.0 & 174.0 \\
	\textit{Mansonia} & \textit{uniformis} & 3.4 & 4.0 & 4.4 & 4.9 & 5.5 & 4.5 & 0.7 \\
	\textit{Aedes} & \textit{polynesiensis} & 3.2 & 3.7 & 4.1 & 4.5 & 5.1 & 4.1 & 0.6 \\
	\textit{Anopheles} & \textit{rivulorum} & 2.8 & 3.4 & 3.8 & 4.2 & 5.0 & 3.9 & 0.7 \\
	\textit{Anopheles} & \textit{melas} & 2.8 & 3.3 & 3.7 & 4.1 & 4.8 & 3.8 & 0.7 \\
	\textit{Anopheles} & \textit{culicifacies s.l.} & 2.8 & 3.3 & 3.7 & 4.1 & 4.7 & 3.7 & 0.6 \\
	\textit{Anopheles} & \textit{subpictus s.l.} & 2.2 & 3.1 & 3.7 & 4.5 & 7.0 & 4.3 & 6.6 \\
	\textit{Aedes} & \textit{samoanus} & 2.7 & 3.2 & 3.6 & 4.0 & 4.9 & 3.7 & 1.0 \\
	\textit{Anopheles} & \textit{quadrimaculatus} & 2.4 & 3.1 & 3.5 & 4.0 & 4.8 & 3.6 & 0.8 \\
	\textit{Anopheles} & \textit{stephensi} & 2.1 & 2.9 & 3.5 & 4.2 & 5.7 & 3.7 & 1.5 \\
	\textit{Anopheles} & \textit{darlingi} & 2.3 & 2.9 & 3.4 & 3.8 & 4.7 & 3.4 & 0.9 \\
	\textit{Aedes} & \textit{sollicitans} & 1.6 & 2.3 & 2.9 & 3.6 & 6.1 & 3.3 & 2.5 \\
	\textit{Aedes} & \textit{vexans} & 1.7 & 2.2 & 2.7 & 3.1 & 4.2 & 3.0 & 3.3 \\
	\textit{Anopheles} & \textit{cruzii} & 1.2 & 1.6 & 2.0 & 2.5 & 4.0 & 2.2 & 1.3 \\
	\textit{Anopheles} & \textit{bellator} & 1.0 & 1.4 & 1.7 & 2.1 & 3.6 & 2.1 & 5.4 \\
		\hline
		\textit{Culex} & \textit{} & 2.5 & 4.0 & 5.4 & 7.6 & 12.3 & 6.2 & 3.2 \\
		\textit{Anopheles} & \textit{} & 2.1 & 3.6 & 4.9 & 6.8 & 10.8 & 5.5 & 2.9 \\
		\textit{Mansonia} & \textit{} & 3.3 & 4.0 & 4.5 & 4.9 & 5.6 & 4.5 & 0.8 \\
		\textit{Aedes} & \textit{} & 2.5 & 3.3 & 3.8 & 4.3 & 5.2 & 3.8 & 0.9 \\
		\hline
		\textit{Overall} & \textit{} & 2.0 & 3.3 & 4.7 & 6.4 & 10.2 & 5.2 & 2.6 \\
	\end{tabular}
	\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from dissection studies.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std. dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
	\label{tab:dissection_estimated_lifespans_chron}%
\end{table}

\section{Detinova parity analysis}

\begin{table}[htbp!]
\begin{tabular}{l|l|l|l|l|l|l|l|l}
	\textbf{Genus} & \textbf{Species} & \textbf{5$\%$} & \textbf{25$\%$} & \textbf{50$\%$} &
	\textbf{75$\%$} & \textbf{95$\%$} & \textbf{Mean} & \textbf{Std. dev.} \\
	\hline
	\textit{Anopheles} & \textit{funestus} & 5.1 & 8.3 & 11.9 & 18.2 & 41.5 & 16.7 & 20.8 \\
	\textit{Anopheles} & \textit{funestus s.l.} & 4.4 & 7.8 & 13.2 & 24.0 & 75.5 & 25.5 & 58.3
	\\
	\textit{Anopheles} & \textit{nili s.l.} & 5.6 & 8.9 & 12.4 & 17.7 & 36.1 & 16.0 & 17.0 \\
	\textit{Anopheles} & \textit{gambiae} & 4.0 & 6.1 & 8.3 & 12.0 & 22.1 & 10.3 & 8.5 \\
	\textit{Anopheles} & \textit{gambiae (Form M)} & 3.5 & 5.6 & 8.2 & 12.4 & 27.4 & 11.2 & 14.0
	\\
	\textit{Anopheles} & \textit{arabiensis} & 3.2 & 4.9 & 7.2 & 11.2 & 24.0 & 9.8 & 10.0 \\
	\textit{Anopheles} & \textit{gambiae s.l.} & 3.4 & 5.7 & 8.8 & 14.4 & 35.7 & 13.3 & 19.7 \\
	\textit{Anopheles} & \textit{leucosphyrus} & 5.0 & 8.1 & 11.0 & 15.4 & 29.0 & 13.8 & 26.4 \\
	\textit{Anopheles} & \textit{leucosphyrus s.l.} & 5.0 & 7.9 & 11.0 & 15.8 & 29.7 & 14.1 &
	25.3 \\
	\textit{Anopheles} & \textit{aconitus} & 3.9 & 6.6 & 9.5 & 13.7 & 29.4 & 13.4 & 65.1 \\
	\textit{Anopheles} & \textit{dirus (Sp. A)} & 4.9 & 7.4 & 10.3 & 14.9 & 30.2 & 13.1
	& 10.7 \\
	\textit{Anopheles} & \textit{dirus s.l.} & 4.3 & 6.6 & 9.4 & 14.3 & 28.8 & 12.3 & 10.5 \\
	\textit{Anopheles} & \textit{minimus (Sp. A)} & 4.1 & 6.9 & 10.2 & 16.0 & 37.2 &
	15.0 & 24.5 \\
	\textit{Anopheles} & \textit{minimus s.l.} & 3.9 & 5.8 & 7.9 & 11.2 & 20.5 & 9.6 & 6.7 \\
	\textit{Anopheles} & \textit{barbirostris s.l.} & 3.8 & 5.9 & 7.9 & 11.2 & 23.1 & 10.4 &
	11.9 \\
	\textit{Anopheles} & \textit{maculatus} & 3.2 & 5.0 & 7.0 & 10.5 & 21.7 & 9.2 & 8.8 \\
	\textit{Anopheles} & \textit{maculatus s.l.} & 3.4 & 5.3 & 7.6 & 11.7 & 25.6 & 10.3 & 10.5
	\\
	\textit{Anopheles} & \textit{sundaicus s.l.} & 3.2 & 4.9 & 6.9 & 10.6 & 22.4 & 9.3 & 9.3 \\
	\textit{Anopheles} & \textit{subpictus s.l.} & 2.9 & 4.6 & 6.8 & 11.0 & 25.6 & 10.0 & 13.6
	\\
	\textit{Anopheles} & \textit{anthropophagus (lesteri)} & 3.2 & 4.7 & 6.4 & 9.1 & 18.1 & 8.1
	& 9.3 \\
	\textit{Anopheles} & \textit{annularis s.l.} & 4.3 & 5.4 & 6.3 & 7.4 & 10.1 & 6.7 & 2.1 \\
	\textit{Anopheles} & \textit{sinensis} & 3.5 & 4.8 & 6.2 & 8.2 & 13.1 & 7.0 & 3.6 \\
	\textit{Anopheles} & \textit{fluviatilis s.l.} & 3.6 & 4.8 & 6.0 & 7.5 & 11.0 & 6.5 & 2.6 \\
	\textit{Anopheles} & \textit{culicifacies s.l.} & 2.6 & 4.1 & 5.9 & 9.4 & 20.1 & 8.3 & 10.2
	\\
	\textit{Anopheles} & \textit{farauti (No.1)} & 3.6 & 4.7 & 5.5 & 6.6 & 9.3 & 5.9 &
	2.1 \\
	\textit{Anopheles} & \textit{farauti s.l.} & 3.9 & 4.9 & 5.6 & 6.6 & 8.7 & 5.9 & 1.6 \\
	\textit{Anopheles} & \textit{darlingi} & 3.1 & 4.9 & 7.4 & 12.0 & 27.6 & 10.8 & 13.4 \\
	\textit{Anopheles} & \textit{pseudopunctipennis s.l.} & 3.9 & 5.4 & 6.5 & 8.0 & 13.0 & 7.4 &
	4.5 \\
	\textit{Anopheles} & \textit{aquasalis} & 3.0 & 4.6 & 6.5 & 9.7 & 20.0 & 8.5 & 8.0 \\
	\textit{Anopheles} & \textit{albitarsis (Sp C)} & 5.2 & 9.5 & 16.4 & 32.6
	& \text{$>$100} & 66.8 & \text{$>$100} \\
	\textit{Anopheles} & \textit{albitarsis (Sp A)} & 2.3 & 3.3 & 4.3 & 5.9 & 10.1 &
	5.1 & 3.3 \\
	\textit{Anopheles} & \textit{albitarsis (Sp. B)} & 2.2 & 3.1 & 4.1 & 5.4 & 9.0 & 4.7 & 2.7
	\\
	\textit{Anopheles} & \textit{albitarsis s.l.} & 2.6 & 4.0 & 5.8 & 9.2 & 19.9 & 8.2 & 9.5 \\
	\textit{Anopheles} & \textit{albimanus} & 2.7 & 4.1 & 5.5 & 8.0 & 15.0 & 6.8 & 4.8 \\
	\textit{Anopheles} & \textit{nuneztovari s.l.} & 2.6 & 3.6 & 4.5 & 6.1 & 11.4 & 5.5 & 3.8 \\
\end{tabular}
\caption{\textbf{Posterior summaries of mean female mosquito lifespan in terms of chronological time from Detinova-type dissection.} The 5\%, 25\%, 50\%, 75\% and 95\% columns indicate the respective quantiles of the posterior distribution for mean lifespan, and the `Mean' and `Std. dev.' columns indicate the posterior mean and standard deviation of mean lifespan.}
\label{tab:detinova_lifespan_chronological}%
\end{table}

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mcPowerAnalysis_senescence.pdf}}
	\caption{\textbf{MRR: statistical power analysis for senescence detection.} The plot shows the statistical power to detect senescence for A. three case study populations as a function of B. study length and C. number released. For each parameter set we generated 500 simulated data series using the negative binomial sampling model with a gompertz hazard function ($\lambda(t) = \alpha e^{\beta t}$), and estimated the $\beta$ parameter using maximum likelihood and tested the null hypothesis $\beta=0$ (constant mortality risk) against the alternative $\beta>0$ using an likelihood ratio test with a 5\% test size. The error bars show the standard deviation in the power for each parameter set. The recapture probability parameters ($\psi$ and $\kappa$) used were the averages of those estimated for the actual data.}
	\label{fig:mrr_mcPowerAnalysis_senescence}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/elpd_vs_exponential_ordered_dissection.png}}
	\caption{\textbf{Polovodova dissection: evidence for senescence.}  Here, we consider the predictive accuracy of age-dependent models of mortality versus the exponential model by species. The predictive accuracy was determined using K-Fold cross-validation as described in SOM; the measure of accuracy presented here by the central dots is the difference in estimated expected log pointwise predictive density compared to the exponential model. The lower and upper whiskers represent the lower and upper bounds of the 95\% confidence interval in predictive accuracy. The species have been ordered so that those with the highest average predictive accuracy across all age-dependent models compared to the exponential model appear at the top. If all age-dependent models outperform the exponential, we deem this as evidence for age-dependent mortality (`+'); if a subsample of models perform better, we deem the evidence ambiguous (`?'); and if no age-dependent model exceeds the performance of the exponential, we conclude no evidence in favour of senescence (`-').}
	\label{fig:dissection_senescence}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/elpd_comparison.pdf}}
	\caption{\textbf{Evidence for senescence: comparing evidence from Gompertz model across MRR and Polovodova dissection analyses.} The points indicate the estimated predictive accuracy compared with the exponential model (see SOM Methods) and the lower and upper whiskers represent 2.5\% and 97.5\% quantiles on this quantity.}\label{fig:elpd_comparison}
\end{figure}
